---
title: "student performance2.0"
author: "Alex Constantinescu & Matei Grosu"
date: "1/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Introduction

We have taken a data set on the student performance on exams regarding math,reading and writing scores.
The data set is taken from  https://www.kaggle.com/spscientist/students-performance-in-exams
We have searched for a data set that is already cleaned.
The data set provides information on the student performance at reading writing and math exams together with data about gender, race,parental education,lunch and test preparation . In this report we will explore the data and analyse it.Specifically, we'll use data visualisation and transformation to explore the choosen data set.

Question:
1.What type of variation occurs within the variables? 

2.What type of covariation occurs between the variables?


## Import data

First, we have to import the data set into the rStudio environment so that we can start explore the data and analyse it.
```{r}
stPerform <- read_csv("/Users/alex/Desktop/UNIVERSITY/SEM3/DASC/StudentsPerformance.csv")
stPerform
```
## Displaying the internal structure of the data frame:
```{r}
str(stPerform)
```

## Some descriptive information of the data:
```{r}
summary(stPerform)
```
From the descriptive information of the data we understood of what types are the data's variables. We have continous variables(math score, reading score and writing score) and categorical variables(gender,race/ethnicity, parental level of education, lunch, test preparation). Also, we want to search for missing values and than to rename the column names to be easier to work with

## Trying to tidy the data
The column names are not that good to work with so we decided to rename them in order to be easier to work with.
```{r}
colnames(stPerform)<-c("Gender","Race","Parental_Education","Lunch","Test_Prep","Math_Score","Reading_Score","Writing_Score")
head(stPerform)
```
We want to find out if there are any missing values to deal with
Dealing with missing values:


Count missing values:
```{r}
sum(is.na(stPerform))
```
We do not have missing values.



## Let's create a new column for the average score at all subjects

```{r}
stPerformAvg <- mutate(stPerform,
       avgScore = (Math_Score+Reading_Score+Writing_Score)/3)
stPerformAvg
```
## Data Viualization:

# we want to see how females did in there exams, let's create a new table only with the females:
```{r}
fPerform <-filter(stPerformAvg,Gender=="female")
fPerform
```
# let's see the distribution of scores for females
```{r}
ggplot(data = fPerform) +
  geom_histogram(mapping = aes(x = Math_Score), binwidth = 0.5)
```
## Let's do the same thing for the males,so let's create a table only with males
```{r}
mPerform <-filter(stPerformAvg,Gender=="male")
mPerform
```

# let's see the distribution of scores for males
```{r}
ggplot(data = mPerform) +
  geom_histogram(mapping = aes(x = Math_Score), binwidth = 0.5)
```

```{r}
ggplot(data = stPerformAvg, mapping = aes(x = Math_Score, colour = Gender)) +
  geom_freqpoly(binwidth = 10)
```
Creating a ggplot to visualize the relationship between math scores and average scores related to gender. 
```{r}
ggplot(data = stPerformAvg) +
  geom_point(mapping = aes(x =Math_Score , y =avgScore, color=Gender))
```
We observe that females tend to have higher scores that males. There is a pozitive relationship between reading and math score

More plots:
```{r}
ggplot(data = stPerform) +
  geom_point(mapping = aes(x =Math_Score , y =Reading_Score, color=Race))
```
This plot does not help us too much as it is not that clear the distribution of race groups.

Let's try to visualize the distribution of the races 
```{r}
ggplot(data = stPerform) +
  geom_bar(mapping = aes(x = Race))
```
We observe that the majority of people is from the group c.


## Let's see how the parental educatian affects the scores of the students

```{r}
ggplot(data = stPerformAvg) +
  geom_point(mapping = aes(x =Math_Score , y =avgScore, color=Gender ))+
  facet_wrap(~ Parental_Education,nrow = 2)
```

## Let's see the distribution of average scores:
```{r}
ggplot(data = stPerformAvg) +
  geom_histogram(mapping = aes(x = avgScore), binwidth = 0.5)
```


```{r}
ggplot(data = stPerform) +
  geom_point(mapping = aes(x =Math_Score , y =Reading_Score, color=Lunch))
```

We see that the students that had a lunch before exams did better
Students performance related to parents education:
We assume that higher the education parents have higher the students grades are

## Plot exam scores by gender to determinate if there is a different score tendency for each gender

#Math scores by gender: 

```{r}
pl <- ggplot(stPerformAvg, aes(Math_Score)) + geom_histogram(binwidth=5, color="gray", aes(fill=Gender))
pl <- pl + xlab("Math Scores") + ylab("Gender") + ggtitle("Math Scores by Gender")
pl

```
#Reading scores by gender: 
```{r}
pl1 <- ggplot(stPerformAvg, aes(Reading_Score)) + geom_histogram(binwidth=5, color="gray", aes(fill=Gender))
pl1 <- pl1 + xlab("Reading Scores") + ylab("Gender") + ggtitle("Reading Scores by Gender")
pl1
```
# Writing scores by gender:
```{r}
pl2 <- ggplot(stPerformAvg, aes(Writing_Score)) +  geom_histogram(binwidth=5, color="gray", aes(fill=Gender))
pl2 <- pl2 + xlab("Writing Scores") + ylab("Gender") + ggtitle("Writing Scores by Gender")
pl2
```

# Average score by gender:
```{r}
pl3 <- ggplot(stPerformAvg, aes(avgScore)) +  geom_histogram(binwidth=5, color="gray", aes(fill=Gender))
pl3 <- pl2 + xlab("Average Scores") + ylab("Gender") + ggtitle("Average Scores by Gender")
pl3
```
## Let's createcreate some boxplots in order to provide some measures of central tendency and visual of outliers:
#Boxplot of writing scores and test preparation by gender:
```{r}

bx <- ggplot(stPerformAvg, aes(Gender, Writing_Score, color = Test_Prep))
bx <- bx + geom_boxplot()
bx <- bx + ggtitle("Writing scores by Gender Boxplot")
bx <- bx + xlab("Gender") + ylab("Writing Scores")
bx
```
#Boxplot of Math scores and test preparation by gender:
```{r}
bx1 <- ggplot(stPerformAvg, aes(Gender, Math_Score, color = Test_Prep))
bx1 <- bx1 + geom_boxplot()
bx1 <- bx1 + ggtitle("Math scores by Gender Boxplot")
bx1 <- bx1 + xlab("Gender") + ylab("Math Scores")
bx1
```
#Boxplot of Reading scores and test preparation by gender:
```{r}
bx2 <- ggplot(stPerformAvg, aes(Gender, Reading_Score, color = Test_Prep))
bx2 <- bx2 + geom_boxplot()
bx2 <- bx2 + ggtitle("Reading scores by Gender Boxplot")
bx2 <- bx2 + xlab("Gender") + ylab("Reading Scores")
bx2
```
#Boxplot of Average scores and test preparation by gender:
```{r}
bx3 <- ggplot(stPerformAvg, aes(Gender, avgScore, color = Test_Prep))
bx3 <- bx3 + geom_boxplot()
bx3 <- bx3 + ggtitle("Average scores by Gender Boxplot")
bx3 <- bx3 + xlab("Gender") + ylab("Average Scores")
bx3
```

After visualizing and analizing the boxplots we observed that students who completed the test preparation has better scores in all three exams.

Moreover, it is shown that male students has better scores in math, while female students have in reading and writing.

There are outliers in all three exams.

In average female students did better than male students, and in average students who completed the test preparation did better than the ones who did not.



## Linear regression(Alex Constantinescu):

```{r}
linearmodel <- lm(Math_Score~avgScore, data= stPerformAvg)
summary(linearmodel)
```
## Interpretation of the linear regression output:
```{r}
plot(linearmodel)
```


##Linear regression model: assumptions
A linear regression makes certain assumptions about the 
data, which you must check:

1. Homoscedasticity of errors: the variance is homogenous

2. Normality: Errors are normally distributed

3. Independence of observations: data was collected using 
valid sampling methods, there are no hidden biases or 
relationships 

4. The relationship between X and Y is linear (i.e. the line is 
straight, not curved)

## Residual analysis:

#1.Residual vs Fitted: Linearity assumption

This plot shows whether there are non-linear patterns.
The red line follows the doted one so it is a perfect scenario.
The residuals appear to be equally spread.

#2.Q-Q(quantitive-quantile):Normality assumption
This plot shows if the residuals are normally distributed. If they are, they follow the straight line in this plot.
We see here that apart from some outliers, the values fall on or around the straight line

#3.  Scale-Location: Homoscedasticity assumption:

This plot tests for homoscedasticity, i.e. if the residuals have equal variance along the regression line.

In this plot, we have a lot of cases so the variance is equlaity spreaded among the plot.

#4. Residuals vs Leverage


This plot is used to spot influential cases in the model.























